### 2025-03-22 paper/paper.tex
add tex files

\documentclass[letterpaper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{bussproofs}

\newgeometry{vmargin={18mm}, hmargin={20mm}}

\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\mtt}[1]{\mathtt{#1}}

\begin{document}

\end{document}