### 2025-04-16 paper/paper.tex
introduction

\documentclass[letterpaper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{bussproofs}

\newgeometry{vmargin={18mm}, hmargin={20mm}}

\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\mtt}[1]{\mathtt{#1}}

\title{borrow-lambda}
\author{Taran Dwivedula, Jeff Zhang}

\begin{document}

\maketitle

\begin{abstract}
    Borrow checking and ownership is a solution to many of the problems that commonly arise in programming languages with manual memory management like C.
    Like the strategy of garbage collection implemented in several common general purpose programming languages, borrow checking eliminates the
    need for programmers to manually allocate and free the memory their program uses, which is the source of many human errors in programming.
    However, the need for high performance code makes garbage collection a sub-optimal solution, as garbage collection introduces performance
    overheads, non-determinism, and potentially memory fragmentation. Borrow checking and ownership is an alternative that takes a different
    approach - it relies on the strength of the programming language's typesystem. It is a set of rules that allows the compiler to
    automatically determine where to allocate and free memory, both eliminating the need of a garbage collector and the need for a programmer
    to manually allocate and free memory. In this paper, we discuss borrow checking as a solution to common memory errors, and provide
    and implementation on a small extension of the simply-typed lambda calculus.
\end{abstract}

\section{Introduction}
Systems-level programming mandates high performance code; that is, code that runs with minimal overhead, and uses memory efficiently. Minimizing
fragmentation on the heap is essential, especially for servers and embedded systems. It's why a programming language like C has been very popular
in this domain for decades. C is an abstraction directly from assembly that was
designed to make code easier to read and write. As such, control of heap allocated
memory remains fully with the programmer. The practice of writing good and safe code
is left to the programmer - the idea being, a skilled programmer would be able to write
memory safe code without mistakes, and just like assembly, mistakes would not result
in errors - just undefined behaviour.

Over time, we have seen that relying on programmers to be skilled enough to avoid
making memory errors in their code is not feasible. When codebases become large scale,
these errors become extremely hard to avoid. Computer viruses and exploits due to
things like buffer overflows affect many people and are still present in a large
amount of software today.

For general purpose programming, the concept of a garbage collector was able to solve
this problem. A garbage collector finds areas of memory that are no longer referenced
at some point in the program, and frees the memory. Different strategies exist, such
as reference counting, stop and copy, and mark and sweep strategies. Usually, the
garbage collector is required to run as a separate process, adding performance
overhead. Some strategies of garbage collection allow the garbage collector to run
concurrent to the program, but fragment the memory. Other strategies avoid memory
fragmentation, but require the running program to be paused so the memory can be
copied over in a non-fragmented fashion. In any case, garbage collection results
in the overhead of another process, as well as sometimes non-determinism in the layout
of memory, both of which are extremely sub-optimal for high performance systems code.

Work has been done to improve manually memory managed languages like C in terms of
safety, by strenghthening the typesystem itself, and attacking the semantics of the
language. Cyclone, a Cornell lab project, developed a dialect of C that works to
ensure memory safety at compile time. Among the features included were statically
checked lifetimes of pointer variables, and restrictions on types of pointers which
work to ensure safety with regards to multi-aliasing (add citation). Similarly, the
concept of ownership types was conceived as a software design principle, particularly
in research on concurrent programming, where an object governs an area of memory,
and administers any modifications to it (add citations). In the modern day,
many of these features were materialized in the Rust programming language, a popular
systems-level programming language that is making the case for the assistance of
borrow checking and ownership for high performance code.

In this paper, we implement borrow checking and ownership on a small extension of the
simply typed lambda calculus. The paper is organized as follows.

\end{document}