### 2025-04-17 paper/paper.tex
overview

\documentclass[letterpaper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{bussproofs}
\usepackage{listings}

\newgeometry{vmargin={18mm}, hmargin={20mm}}

\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\mtt}[1]{\mathtt{#1}}

\title{borrow-lambda}
\author{Taran Dwivedula, Jeff Zhang}

\begin{document}

\maketitle

\begin{abstract}
    Borrow checking and ownership is a solution to many of the problems that commonly arise in programming languages with manual memory management like C.
    Like the strategy of garbage collection implemented in several common general purpose programming languages, borrow checking eliminates the
    need for programmers to manually allocate and free the memory their program uses, which is the source of many human errors in programming.
    However, the need for high performance code makes garbage collection a sub-optimal solution, as garbage collection introduces performance
    overheads, non-determinism, and potentially memory fragmentation. Borrow checking and ownership is an alternative that takes a different
    approach - it relies on the strength of the programming language's typesystem. It is a set of rules that allows the compiler to
    automatically determine where to allocate and free memory, both eliminating the need of a garbage collector and the need for a programmer
    to manually allocate and free memory. In this paper, we discuss borrow checking as a solution to common memory errors, and provide
    and implementation on a small extension of the simply-typed lambda calculus.
\end{abstract}

\section{Introduction}
Systems-level programming mandates high performance code; that is, code that runs with minimal overhead, and uses memory efficiently. Minimizing
fragmentation on the heap is essential, especially for servers and embedded systems. It's why a programming language like C has been very popular
in this domain for decades. C is an abstraction directly from assembly that was
designed to make code easier to read and write. As such, control of heap allocated
memory remains fully with the programmer. The practice of writing good and safe code
is left to the programmer - the idea being, a skilled programmer would be able to write
memory safe code without mistakes, and just like assembly, mistakes would not result
in errors - just undefined behaviour.

Over time, we have seen that relying on programmers to be skilled enough to avoid
making memory errors in their code is not feasible. When codebases become large scale,
these errors become extremely hard to avoid. Computer viruses and exploits due to
things like buffer overflows affect many people and are still present in a large
amount of software today.

For general purpose programming, the concept of a garbage collector was able to solve
this problem. A garbage collector finds areas of memory that are no longer referenced
at some point in the program, and frees the memory. Different strategies exist, such
as reference counting, stop and copy, and mark and sweep strategies. Usually, the
garbage collector is required to run as a separate process, adding performance
overhead. Some strategies of garbage collection allow the garbage collector to run
concurrent to the program, but fragment the memory. Other strategies avoid memory
fragmentation, but require the running program to be paused so the memory can be
copied over in a non-fragmented fashion. In any case, garbage collection results
in the overhead of another process, as well as sometimes non-determinism in the layout
of memory, both of which are extremely sub-optimal for high performance systems code.

Work has been done to improve manually memory managed languages like C in terms of
safety, by strenghthening the typesystem itself, and attacking the semantics of the
language. Cyclone, a Cornell lab project, developed a dialect of C that works to
ensure memory safety at compile time. Among the features included were statically
checked lifetimes of pointer variables, and restrictions on types of pointers which
work to ensure safety with regards to multi-aliasing (add citation). Similarly, the
concept of ownership types was conceived as a software design principle, particularly
in research on concurrent programming, where an object governs an area of memory,
and administers any modifications to it (add citations). In the modern day,
many of these features were materialized in the Rust programming language, a popular
systems-level programming language that is making the case for the assistance of
borrow checking and ownership for high performance code.

In this paper, we implement borrow checking and ownership on a small extension of the
simply typed lambda calculus. The paper is organized as follows. Section 2 provides an
overview of the components of the borrow checking and ownership strategy. Section 3
discusses our implementation of a borrow checker in \texttt{borrow-lambda}.

\section{Overview}
The borrow checking and ownership strategy can be divided into three separate components
that work together - Ownership, Borrowing, and Lifetimes. This section will conceptually
describe the three. As the Rust programming language is the most well known widespread
implementation of these concepts, we will reference it for illustrative purposes.

\subsection{Ownership}
Rust implements ownership as described with three main rules (add citation).
\begin{itemize}
    \item Each value in Rust has an owner.
    \item There can only be one owner at a time.
    \item When the owner goes out of scope, the value will be dropped.
\end{itemize}
In reality, this is an instance of an \textit{affine type system} - an instance
of a substructural type system based on substructural logic where the rule
of contraction is not permitted. In code, this corresponds to the language's
move semantics - if a value isn't given the ability to be copied, it is instead
\textit{moved}. Ownership and scope of the value is transferred to the new binder.
This allows the compiler to statically determine where the memory of the value
should be freed, and where the value, as well as names tied to it, are allowed to
be used, absolving this responsibility from the programmer.
\begin{figure}[h]
    \begin{lstlisting}
        fn main() {
            let nums = vec![1, 2, 3];

            do_something_with_vec(nums);
            do_something_with_vec(nums);  // error
        }

        fn do_something_with_vec(v: Vec<u32>) {}
    \end{lstlisting}

    \caption{Example Rust code illustrating move semantics}
    \label{ms}
\end{figure}
Figure \ref{ms} illustrates an example. We create a vector of unsigned
integers - an object whose type does not permit copying - and bind it to the
variable \texttt{nums}. We say the variable \texttt{nums} \textit{owns} this
vector and its associated memory. Up to this point, we should expect \texttt{nums}
to go out of scope at the end of \texttt{main}, and the compiler would automatically
free the memory at this time. However, we instead pass \texttt{nums} as a parameter
into the function \texttt{do\_something\_with\_vec}. Then, as previously described,
the vector gets \textit{moved} into \texttt{do\_something\_with\_vec}. The vector is
no longer bound to the name \texttt{nums}, which causes the second call to
\texttt{do\_something\_with\_vec} to fail. Additionally, as the new owner, the
function call administers the scope of this vector as well. As such, the vector's
memory gets freed at the end of the function call. At this point, all access points to this memory are unusable and trigger compile time errors, which is what we want.

What if we did actually want to use the same vector twice inside of \texttt{main}?
Currently, it is quite inconvenient that the vector gets detached from the \texttt{nums}
variable after the first use of it. We may rewrite the code such that
\texttt{do\_something\_with\_vec} returns the vector, passing ownership of it back
to the caller, allowing us to re-bind the vector to the \texttt{nums} variable.
That is one possible solution, but what we really want is to provide some sort
of reference, rather than temporarily giving ownership and taking it back.
Crucially, it eliminates one nice guarantee we can make, which is the state
of the value before and after we lend it. It would be nice to know, at a type level,
whether the value we lent was mutated before we received it back. This is made
possible with borrowing, which we describe in the next subsection.

\subsection{Borrowing}
Borrowing is a form of restricted pointers that allows us to give access of a value
in memory, without relinquishing ownership of the value. Access is given through
references to the value. These references are set to be either immutable or mutable
by the programmer, meaning that references may only provide a view on the data
without providing permission to mutate the data. Enforcement is an easy check during
compile time.

The concept of mutable and immutable references is simple enough, but what makes them
useful is the rules that determine when they are valid, and how they affect the
access points of memory. Particularly, the following rules related to references
are enforced at compile time.
\begin{itemize}
    \item At any given time, a value can have either one mutable reference pointing to it, or any amount of immutable references pointing to it, but not both.
    \item The owner of a value cannot mutate or give up ownership of a value while any references to the value are in scope.
\end{itemize}
Again, both easy checks at compile time. It is easily seen that these two restrictions
ensure two important conditions, that is, only one object is able to mutate a value
at a time, and that values will not be mutated or freed while being read from. In
other words, this eliminates the problem of mutable aliasing.

\subsection{Lifetimes}
It is important to know how long a value is valid for, before its associated memory
is freed or repurposed. Being able to check this at compile time is a large component
of creating a memory safe language. The aformentioned Cyclone first experimented
with this by adding regions to C, and it is the basis of Rust's lifetimes feature.
A potential, and extremely common error, is that we may have a pointer to some value
that we continue to use even after the value's associated memory gets freed - a dangling pointer.

With the groundwork of ownership and borrowing laid out, the solution to this is
simple - when we borrow a value, we must ensure that the value's associated memory
does not get freed while a reference to it is still in scope. Or rather, we should
not allow a reference to be created if its scope exceeds the window for which
the value's memory remains intact. We formalize this with lifetimes.

Generally, the \textit{lifetime} of a value, or the window in which its memory
can be used before it is freed, is taken from the scope of its owner. These
lifetimes are embedded into the typesystem as any references are additionally
tagged with the lifetime of the value it is pointing to. Then, these references
are checked with a subtyping relation, where a reference with a larger scope
is a subtype of a reference with a smaller scope. This allows us to check that
all borrows, or references, go out of scope before the value that they are borrowing
is freed.

\texttt{TODO: maybe add example code, explanation for functions polymorphic over lifetimes}

\section{Implementation}

\end{document}